---
title: "Bar plots"
author: "Ashwin Malshé"
date: "21 September 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Bar graph

As seen previously, bar graphs are used when we have a categorical variable.


What happens when you have already created summary statistics for making a bar plot?

```{r}
mpgSum <- mpg %>%
  group_by(manufacturer) %>% 
  summarise(hwy_mean = mean(hwy),
            hwy_n    = n())

head(mpgSum)
```

Can we now use the summarized data to plot hwy_n on y axis?

```{r}
ggplot(mpgSum, aes(x = manufacturer, y = hwy_n)) +
  geom_bar()

```

Correct the plot above

```{r}
ggplot(mpgSum, aes(x = manufacturer, y = hwy_n)) +
  geom_bar(stat = "identity")
```

OR you could also use a different `geom`

```{r}
ggplot(mpgSum, aes(x = manufacturer, y = hwy_n)) +
  geom_col()
```


Now with the new tool, you can also create bar plot with *any* Y axis variable you want

```{r}
ggplot(mpgSum, aes(x = manufacturer, y = hwy_mean)) +
  geom_col()

```

If you don't want to create an intermediate data set, use magrittr pipe

```{r}
mpg %>%
  group_by(manufacturer) %>% 
  summarise(hwy_mean = mean(hwy),
            hwy_n    = n()) %>% 
  ggplot(aes(x = manufacturer, y = hwy_mean)) +
    geom_col()

```

In the above bar graph, we can introduce colors for `drv` which is `c("4", "f", "r")`

```{r}
mpg %>%
  group_by(manufacturer, drv) %>% 
  summarise(hwy_mean = mean(hwy),
            hwy_n    = n()) %>% 
  ggplot(aes(x = manufacturer, y = hwy_mean)) +
    geom_col(aes(fill = drv))
```


Remove the legend

```{r}
mpg %>%
  group_by(manufacturer, drv) %>% 
  summarise(hwy_mean = mean(hwy),
            hwy_n    = n()) %>% 
  ggplot(aes(x = manufacturer, y = hwy_mean)) +
    geom_col(aes(fill = drv), show.legend = FALSE)

```

Add labels to the bars

```{r}
mpg %>%
  group_by(manufacturer, drv) %>% 
  summarise(hwy_mean = mean(hwy),
            hwy_n    = n()) %>% 
  ggplot(aes(x = manufacturer, y = hwy_mean)) +
  geom_col(aes(fill = drv)) +
  geom_text(aes(label = round(hwy_mean, 1),
                group = drv),
                position = position_stack(vjust = 0.5))
```

