---
title: "Basic Plots Part 1"
author: "Ashwin Malshé"
date: "21 September 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
theme_set(theme_minimal())
```

The objective of this exercise is to create a few basic plots that are often used in statistical reporting.

### Boxplot

Boxplots show the distribution of a variable

```{r}
ggplot(mpg, aes(y = hwy)) +
  geom_boxplot() 
```


Usually, boxplots for different levels of a discrete variable are more useful:

```{r}
ggplot(mpg, aes(x = manufacturer, y = hwy)) +
  geom_boxplot()
```

We can change the color of the boxplots

```{r}
ggplot(mpg, aes(x = manufacturer, y = hwy)) +
  geom_boxplot(color = "red")
```

Map `color` to `manufacturer`

```{r}
ggplot(mpg, aes(x = manufacturer, y = hwy, color = manufacturer)) +
  geom_boxplot()
```

```{r}
ggplot(mpg, aes(x = manufacturer, y = hwy, color = "red")) +
  geom_boxplot()
```


**Fill** colors **inside** the boxplot

```{r}
ggplot(mpg, aes(x = manufacturer, y = hwy)) +
  geom_boxplot(color = "red", fill = "blue")
```

Map `fill` to `manufacturer`

```{r}
ggplot(mpg, aes(x = manufacturer, y = hwy, fill = manufacturer)) +
  geom_boxplot()
```

You have a lot of control over what to do with the outliers. Check out the documentation:
https://ggplot2.tidyverse.org/reference/geom_boxplot.html


### Bar graph

The default use of a bar graph is to show the frequency distribution of a categorical variable. In case of `ggplot2`, the y axis is the frequency, which it calculates on its own. Thus, the frequency is a `stat` which `geom_bar()` uses as default Y axis.

```{r}
ggplot(mpg, aes(x = class)) +
  geom_bar()
```


Bar plot on a continuous variable


```{r}
count(mpg, cyl)
```


```{r}
ggplot(mpg, aes(x = cyl)) +
  geom_bar()
```

Fix the problem in the above graph!


```{r}
ggplot(mpg, aes(x = as.character(cyl))) +
  geom_bar()
```

Change color

```{r}
ggplot(mpg, aes(x = manufacturer)) +
  geom_bar(color = "blue")
```

Change the fill

```{r}
ggplot(mpg, aes(x = manufacturer)) +
  geom_bar(color = "blue", fill = "green")
```

Change transparency

```{r}
ggplot(mpg, aes(x = manufacturer)) +
  geom_bar(color = "blue", fill = "green", alpha = 0.5)
```

Where do you get colors names from?

Here is a nice cheatsheet:

https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/colorPaletteCheatsheet.pdf



### Histogram

Histogram is a bar plot for a continuous variable. Making a bar plot for a continuous variable doesn't make sense because the frequency of each unique value is likely to be very low. This leads to mostly a flat bar graph. Histogram solves this problem by creating bins from the continuous variable. It then plots the frequencies of these bins. A common example is to create bins of the ages of people in the sample.

```{r}
ggplot(mpg, aes(x = hwy)) +
  geom_bar()
```


```{r}
ggplot(mpg, aes(x = hwy)) +
  geom_histogram()
```

### Scatterplot

Scatterplot depicts the relationship between two continuous variables. Let's study 

```{r}
ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
```

Add a trend line to the scatterplot

```{r}
ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point() +
  geom_smooth()
```



Adding the title, subtitle, and the caption is really easy.


```{r}
ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point() +
  labs(title = "City Miles vs Highway Miles",
       subtitle = "Based on the data from 1999 and 2008",
       caption = "Created by Ashwin Malshe using mpg dataset.
       mpg dataset is included in ggplot2 package.",
        x = "miles per gallon in city driving",
        y = "miles per gallon in highway driving")
```


We can show subgroups by using various aesthetics. For example, we can color the points based on the drive of the cars.


```{r}
ggplot(mpg, aes(x = cty, y = hwy, color = drv)) +
  geom_point() +
  labs(title = "City Miles vs Highway Miles",
       subtitle = "Based on the data from 1999 and 2008",
       caption = "Created by Ashwin Malshe using mpg dataset.")
```


Finally, instead of showing the groups separately may make sense in some cases. This code will split the data into 3 groups based on the value of `drv` and then create a plot for each subgroup. This is known as a facet plot.


```{r}
ggplot(mpg, aes(x = cty, y = hwy, color = drv)) +
  geom_point() +
  facet_wrap(~ drv)
```

